Sample Based Determination of Angular Observables

### 2014-06-18 draft.tex
Add two comments on combining observables (across P, across q2)

We assume, for a moment, that $P$ describes the distribution of the underlying
physical process truthfully. In that case, all events detected in an experimental setup are in fact, up to
detection efficiencies, distributed as $P$. Application of \refeq{mc-id} then yields
\begin{equation}
    \langle P_i\rangle \to \widehat{\langle P_i\rangle} = \frac{1}{N} \sum_{k=1}^{N} f_i(x^{(k)})\,.
\end{equation}
However, it is often of interest to obtain the $\nuvec$-integrated observables for
certain $\nuvec$ ranges, i.e. binned measurements of $\langle P_i\rangle$\footnote{%
We refer for example to the determination of the angular observables $J_i$ in $\bar{B}\to\bar{K}^*\ell^+\ell^-$
decays, see e.g. \cite{Aaij:2013iag}. For details on the relation between the $P_i$ and the angular observable $J_i$ in the
aforementioned decay, we refer to appendix \ref{app:btokstarll}.
}.
Therefore, we introduce the bin-integrated quantities
\begin{align}
    \langle P_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} P_i(\nuvec) \dd^m \nu\\
    & = \int_{\vec{a}}^{\vec{b}} \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec) \dd^n\theta\right]\dd^m \nu\\
    & = \int \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec)
        \theta^{(m)}(\nuvec - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec) \dd^n\theta\right]\dd^m \nu\,.
\end{align}
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{N} \sum_{k=1}^{N} f_i(x^{(k)}) \theta^{(m)}(\nuvec^{(k)} - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec^{(k)})\,.
\end{equation}
This basically reduces the determination of $P_i$ to a (weighted) counting experiment. As a consequence,
the statistical error on $\what{P_i}$ can be expected to follow a multivariate gaussian distribution \citeneeded.
In this case, the covariance $\cov[\what{P_i},\what{P_j}]$ is of particular interest. The latter can be replaced
by a suitable MC estimator, the sample covariance
\begin{equation}
    \cov[\what{P}_i,\what{P}_j] \to \what{\cov}[\what{P}_i, \what{P}_j]_{\vec{a},\vec{b}}
        = \frac{1}{N - 1} \sum_{k=1}^{N} \big[\hat{f}_i^{(k)} - \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[\hat{f}_j^{(k)} - \widehat{\langle P_j\rangle_{\vec{a},\vec{b}}}\big]
        \theta^{(m)}(\nuvec^{(k)} - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec^{(k)})\,.
\end{equation}

@@

>>>>

We assume, for a moment, that $P$ describes the distribution of the underlying
physical process truthfully. In that case, all events detected in an experimental setup are in fact, up to
detection efficiencies, distributed as $P$. Application of \refeq{mc-id} then yields
\begin{equation}
    \langle P_i\rangle \to \widehat{\langle P_i\rangle} = \frac{1}{N} \sum_{k=1}^{N} f_i(x^{(k)})\,.
\end{equation}
However, it is often of interest to obtain the $\nuvec$-integrated observables for
certain $\nuvec$ ranges, i.e. binned measurements of $\langle P_i\rangle$\footnote{%
We refer for example to the determination of the angular observables $J_i$ in $\bar{B}\to\bar{K}^*\ell^+\ell^-$
decays, see e.g. \cite{Aaij:2013iag}. For details on the relation between the $P_i$ and the angular observable $J_i$ in the
aforementioned decay, we refer to appendix \ref{app:btokstarll}.
}.
Therefore, we introduce the bin-integrated quantities
\begin{align}
    \langle P_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} P_i(\nuvec) \dd^m \nu\\
    & = \int_{\vec{a}}^{\vec{b}} \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec) \dd^n\theta\right]\dd^m \nu\\
    & = \int \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec)
        \theta^{(m)}(\nuvec - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec) \dd^n\theta\right]\dd^m \nu\,.
\end{align}
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{N} \sum_{k=1}^{N} f_i(x^{(k)}) \theta^{(m)}(\nuvec^{(k)} - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec^{(k)})\,.
\end{equation}
This basically reduces the determination of $P_i$ to a (weighted) counting experiment. As a consequence,
the statistical error on $\what{P_i}$ can be expected to follow a multivariate gaussian distribution \citeneeded.
In this case, the covariance $\cov[\what{P_i},\what{P_j}]$ is of particular interest. The latter can be replaced
by a suitable MC estimator, the sample covariance
\begin{equation}
    \cov[\what{P}_i,\what{P}_j] \to \what{\cov}[\what{P}_i, \what{P}_j]_{\vec{a},\vec{b}}
        = \frac{1}{N - 1} \sum_{k=1}^{N} \big[\hat{f}_i^{(k)} - \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[\hat{f}_j^{(k)} - \widehat{\langle P_j\rangle_{\vec{a},\vec{b}}}\big]
        \theta^{(m)}(\nuvec^{(k)} - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec^{(k)})\,.
\end{equation}
We would like to emphasize that the sample covariance should rapidly converge towards the true covariance matrix.
This will allow to construct constraints on optimized observables with correlated uncertainties 
from the experimental constraints on the angular observables. (See \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf}
for definitions of such optimized observables in $B\to K^*\ell^+\ell^-$ decays).

@@

We are aware that setting a bin width that fulfills the aforementioned criterion can be at odds with
other requirements. For instance, in exclusive $b\to s\ell^+\ell^-$ decays, the bin width for
$\vec{\nu} = q^2$ needs to cover a sufficiently large part of the phase space beyond the open charm
threshold. We emphasize that the outcome of an analysis as proposed here can easily combine several
bins by means of a weighted average.\\
